Xor-Some Preregistered Study: Preprocessing

### 2022-04-07 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
exclude subjects resilient to context, remove duplicates

# apply exclusion criteria to main trials
# check range of responses per participant
d_main_fail <- d_main %>% group_by(submission_id) %>%
  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
                                 TRUE ~ TRUE)
         ) %>% filter(passed_main == F)
cat(" Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())

@@

cat(" Nr. of participants left after cleaning: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())
```
[-Exclude the 201st participant for target N=200-]
The exclusion of the last participant isn't necessary anymore, since we only have 200 subjects left upon exclusions due to the initially incorrect native language check.

Write out tidy data:
```{r}
d_full_clean %>% write_csv("./../../data/main/results_prereg_final_tidy.csv")
```

@@

Z-scoring:
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))

d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

@@

Write out z-scored data:
```{r}
d_critical_zScored %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")

>>>>

# apply exclusion criteria to main trials
# check range of responses per participant
d_main_fail <- d_main %>% group_by(submission_id) %>%
  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
                                 TRUE ~ TRUE)
         ) %>% filter(passed_main == F)
# now check for critical trials only
d_main_critical_fail <- d_main %>% group_by(submission_id) %>% 
  filter(condition == "critical") %>%
  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
                                 TRUE ~ TRUE)
         ) %>% filter(passed_main == F)
# now check implicature trials only 
d_main_implicature_fail <- d_main %>% group_by(submission_id) %>% 
  filter((block == "xor") | (block == "some")) %>%
  mutate(passed_main = case_when(max(response) - min(response) <= 10 ~ FALSE,
                                 TRUE ~ TRUE)
         ) %>% filter(passed_main == F)
# check if this failure is the case for certain vignettes more often then for others
d_main_implicature_fail %>% ungroup() %>% group_by(ID, block) %>% count() %>% View()
# --> it seems that the vignettes 28 (xor) and 45 (some) have this issue the most
# check how often the single participants do this
d_main_implicature_fail %>% count() %>% View()
# --> all of them do this on all of the trials - these are 15 people 
# write out a csv with their IDs for an exploratory exclusion
#d_main_implicature_fail %>% count() %>% write_csv("./../../data/main/submission-ids_implicature_fails.csv")

cat(" Subjects providing the same ratings throughout the trials: ", d_main_fail %>% distinct(submission_id) %>% pull() %>% length())

@@

# write out tidy data without the participants providing only 100 ratings on the critical trials
d_clean_exclude_implicatureFails <- anti_join(d_full_clean, d_main_implicature_fail, by = "submission_id")

cat(" Nr. of participants left after cleaning: ", d_full_clean %>% distinct(submission_id) %>% pull() %>% length())

cat(" Nr. of participants left after excluding participants failing implicature manipulations: ", d_clean_exclude_implicatureFails %>% distinct(submission_id) %>% pull() %>% length())

```
[-Exclude the 201st participant for target N=200-]
The exclusion of the last participant isn't necessary anymore, since we only have 200 subjects left upon exclusions due to the initially incorrect native language check.

Write out tidy data:
```{r}
d_full_clean %>% write_csv("./../../data/main/results_prereg_final_tidy.csv")

d_clean_exclude_implicatureFails %>% write_csv("./../../data/main/results_prereg_final_tidy_exclude_implcatureFailures.csv")
```

@@

Z-scoring:
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))
        # response_centered = ifelse(is.na(response_centered), (response-block_mean), response_centered)
d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

# assume that the mistake in the z-scoring is the absence of dropping the duplicates
# or is it the point where the 0 SDs are caught?
d_critical_zScored_noDuplicates <- d_critical_long %>% 
  filter((block_extended == class_condition) | (class_condition =="relevance" & block_extended == "target") )

d_critical_noDuplicates_zScored <- d_critical_zScored_noDuplicates %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))

@@

Write out z-scored data:
```{r}
d_critical_zScored %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
d_critical_zScored_wide %>% write_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
d_critical_noDuplicates_zScored %>% write_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_long.csv")
